---
title: "meow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(igraph)
library(RSiena)
library(ggplot2)
library(network)
library(intergraph)
library(sna)
```

```{r}
vdb.CoreGossip <- load("/Users/joonghochoi/Desktop/Senior thesis work/Data-Josh/CoreGossip/14100Nynke.RData")
```


```{r}
gearyC<-function(res=as.matrix(wellbeing[,2]),meow=friendship[1]){
  #input: data frame of dependent variables
  item<-is.na(res)
  N<-length(rowSums(item))-sum(rowSums(item)>0) #Number of people with no observed values 
  indexes<-c(which(rowSums(item)==0))  #Indexes of ppl with observed values in wellbeing
  Len<-length(indexes)
  #print("length of indexes")
  #print(Len)
  w<-matrix(1:Len**2,nrow=Len,ncol=Len)
  df<-as.data.frame(meow)
  #print("df is ")
  (df)
  #Updating w, which only consists of ppl who filled out wellbeing. 1 if friend, 0 if not. NA if no info.
  for (ii in 1:Len)
  {
    for (jj in 1:Len)
    {
        if (ii==jj)
        {
          w[ii,jj]=0
        }
      else
      {
        i=indexes[ii]
        j=indexes[jj]
        #print("i is ")
        #print(i)
        #print("j is ")
        #print(j)
        #print(df[i,j])
        
        
        
        
        #browser()
        if (df[i,j]==10)
        {
          w[ii,jj]=NA
        }
        if (df[i,j]>4)
        {
          w[ii,jj]=1
        }
        else
        {
          w[ii,jj]=0
        }
        
      }
      
    }
  }
  W<-sum(w)
  (x<-as.data.frame(res))
  numerator_s=0
  denominator_s=0
  x_mean=sum(x,na.rm=TRUE)/(dim(x)[1]-sum(rowSums(is.na(x)))) #correct. Checked manually 
  for (ii in 1:Len){
    x_i<-x[indexes[ii],1]
    for (jj in 1:Len){
      x_j<-x[indexes[jj],1]
      value=w[ii,jj]*(x_i-x_j)**2
      #print(value)
      numerator_s=numerator_s+value
    }
    denominator_s=denominator_s+(x_i-x_mean)**2
  }
  #print("numerator_s")
  #print(numerator_s)
  #print("denominator_s")
  #print(denominator_s)
  answer<-((N-1)*numerator_s)/ (2*W*denominator_s)
  answer
}

```


```{r}
#(a<-as.matrix(c(4,35,25,1))) example from online
(a<-as.matrix(c(2,3,4,5))) #my own example

(b<-matrix(c(0,5,5,0,5,0,0,5,5,0,0,5,0,5,5,0),nrow=4,4))
(c<-as.data.frame(b))
gearyC(a,b) 
```
```{r}
#(a<-as.matrix(c(4,35,25,1))) example from online
(a<-as.matrix(c(2,3,4,5))) #my own example.   Returns matrix,array
item<-list(0,5,5,0,5,0,0,5,5,0,0,5,0,5,5,0)
(b<-matrix(item,nrow=4,4))
(c<-as.data.frame(b))
gearyC(a,b)   #should return 0.75
```


```{r}
#nacf function accepts two inputs: one network and one numeric vector so conversion needed
(a<-as.matrix(c(2,3,4,5))) #my own example.   Returns matrix,array
meow<-c(a)
item<-list(0,1,1,0,1,0,0,1,1,0,0,1,0,1,1,0)
b<-matrix(item,nrow=4,4)
c<-as.data.frame(b)
k<-as.matrix(c)
#k<-as.matrix(b)

k1<-graph_from_adjacency_matrix(k)
k1_1<-asNetwork(k1) #
class(k1_1)

```

```{r}
nacf(k1_1,y=meow,type="geary",mode="digraph")[2]
```

so we know my function works with no missing values.
---

Two code chunks below show nacf doesn't work with na values, whereas mine does. It omits the NA and returns correct Geary C value (proven by manual calculation)
```{r}
#(a<-as.matrix(c(4,35,25,1))) example from online
(a<-as.matrix(c(2,3,4,NA))) #my own example.   Returns matrix,array
item<-list(0,5,5,0,5,0,0,5,5,0,0,5,0,5,5,0)
(b<-matrix(item,nrow=4,4))
(c<-as.data.frame(b))
gearyC(a,b)   
```

```{r}
#nacf function accepts two inputs: one network and one numeric vector so conversion needed
(a<-as.matrix(c(2,3,4,NA))) #my own example.   Returns matrix,array
meow<-c(a)
item<-list(0,1,1,0,1,0,0,1,1,0,0,1,0,1,1,0)
b<-matrix(item,nrow=4,4)
c<-as.data.frame(b)
k<-as.matrix(c)
#k<-as.matrix(b)

k1<-graph_from_adjacency_matrix(k)
k1_1<-asNetwork(k1) #
class(k1_1)
nacf(k1_1,y=meow,type="geary",mode="digraph")[2]
```

------

Now, what if I want to do in mean imputation? 


```{r}
gearyC1<-function(res=wellbeing[,2],meow=f1_1){
  meow1<-mean(res,na.rm=TRUE)
  res[is.na(res)]<-meow1
  #input: data frame of dependent variables
  res=as.matrix(res)
  
  item<-is.na(res)
  N<-length(rowSums(item))-sum(rowSums(item)>0) #Number of people with no observed values 
  indexes<-c(which(rowSums(item)==0))  #Indexes of ppl with observed values in wellbeing
  Len<-length(indexes)
  w<-matrix(1:Len**2,nrow=Len,ncol=Len)
  df<-as.data.frame(meow)
  #print(Len)
  #Updating w, which only consists of ppl who filled out wellbeing. 1 if friend, 0 if not. NA if no info.
  for (ii in 1:Len)
  {
    for (jj in 1:Len)
    {
        if (ii==jj)
        {
          w[ii,jj]=0
        }
      else
      {
        i=indexes[ii]
        j=indexes[jj]
        #print("i is ")
        #print(i)
        #print("j is ")
        #print(j)
        #print(df[i,j])
        #browser()
        if (df[i,j]==10)
        {
          w[ii,jj]=NA
        }
        if (df[i,j]>4)
        {
          w[ii,jj]=1
        }
        else
        {
          w[ii,jj]=0
        }
        
      }
      
    }
  }
  W<-sum(w)
  (x<-as.data.frame(res))
  numerator_s=0
  denominator_s=0
  x_mean=sum(x,na.rm=TRUE)/(dim(x)[1]-sum(rowSums(is.na(x)))) #correct. Checked manually 
  for (ii in 1:Len){
    x_i<-x[indexes[ii],1]
    for (jj in 1:Len){
      x_j<-x[indexes[jj],1]
      value=w[ii,jj]*(x_i-x_j)**2
      #print(value)
      numerator_s=numerator_s+value
    }
    denominator_s=denominator_s+(x_i-x_mean)**2
  }
  #print("numerator_s")
  #print(numerator_s)
  #print("denominator_s")
  #print(denominator_s)
  answer<-((N-1)*numerator_s)/ (2*W*denominator_s)
  answer
}

```

```{r}
(a<-as.matrix(c(2,3,4,3))) #my own example.   Returns matrix,array
item<-list(0,5,5,0,5,0,0,5,5,0,0,5,0,5,5,0)
(b<-matrix(item,nrow=4,4))
(c<-as.data.frame(b))
gearyC1(a,b)   #should return 1.125
```
```{r}
#nacf function accepts two inputs: one network and one numeric vector so conversion needed
(a<-as.matrix(c(2,3,4,3))) #my own example.   Returns matrix,array
meow<-c(a)
item<-list(0,1,1,0,1,0,0,1,1,0,0,1,0,1,1,0)
b<-matrix(item,nrow=4,4)
c<-as.data.frame(b)
k<-as.matrix(c)
#k<-as.matrix(b)

k1<-graph_from_adjacency_matrix(k)
k1_1<-asNetwork(k1) #
class(k1_1)
nacf(k1_1,y=meow,type="geary",mode="digraph")[2]
```

```{r}
(a<-as.matrix(c(2,3,4,NA))) #my own example.   Returns matrix,array
item<-list(0,5,5,0,5,0,0,5,5,0,0,5,0,5,5,0)
(b<-matrix(item,nrow=4,4))
(c<-as.data.frame(b))
gearyC1(a,b)   #should return 1.125
```
#Now testing with the wellbeing data

```{r}
res<-friendshipPreProcess(friendship)
for (ii in 1:6){
  f1<-as.matrix(as.data.frame(res[1])) #wave 1
  f2<-as.matrix(as.data.frame(res[2]))
  f3<-as.matrix(as.data.frame(res[3]))
  f4<-as.matrix(as.data.frame(res[4]))
  f5<-as.matrix(as.data.frame(res[5]))
  f6<-as.matrix(as.data.frame(res[6]))
}

#Use this chunk if don't want to make assumption that loners or ppl who didn't answer survey will remain the same
IdentifyLoners <- function (mat){
  i <- (colSums(f1, na.rm=T) != 0) # T if colSum is not 0, F otherwise
  j <- (rowSums(f1, na.rm=T) != 0) # T if colSum is not 0, F otherwise
  m<-c(i|j)
  m1<-sapply(m, as.numeric)
  m2<- which(m1 %in%0 )
  m2
}
L1<-IdentifyLoners(f1)
L2<-IdentifyLoners(f2)
L3<-IdentifyLoners(f3)
L4<-IdentifyLoners(f4)
L5<-IdentifyLoners(f5)
L6<-IdentifyLoners(f6)
m2<-Reduce(intersect, list(L1,L2,L3))
f1_1<-f1[-m2,-m2]
f2_1<-f2[-m2,-m2]
f3_1<-f3[-m2,-m2]
f4_1<-f4[-m2,-m2]
f5_1<-f5[-m2,-m2]
f6_1<-f6[-m2,-m2]
(m2)
```

```{r}
k78<-wellbeing[,2]
k78<-k78[-m2]
length(k78)
dim(f1_1)
class(a)
class(b)
c<-as.matrix(k78)
class(c)
class(f1_1)
gearyC1(res=as.matrix(k78),meow=f1_1)
```

#removed people who are loners in all six waves from wellbeing. Graph is made from friend network, whic already went through loners removal. Then did mean imputation. Like in the original file

It's cuz f1_1 already got binarized

So now I have a geary C function that can compute with binarized friend network
```{r}
gearyC2<-function(res=wellbeing[,2],meow=f1_1){
  meow1<-mean(res,na.rm=TRUE)
  res[is.na(res)]<-meow1
  #input: data frame of dependent variables
  res=as.matrix(res)
  
  item<-is.na(res)
  N<-length(rowSums(item))-sum(rowSums(item)>0) #Number of people with no observed values 
  indexes<-c(which(rowSums(item)==0))  #Indexes of ppl with observed values in wellbeing
  Len<-length(indexes)
  w<-matrix(1:Len**2,nrow=Len,ncol=Len)
  df<-as.data.frame(meow)
  #print(Len)
  #Updating w, which only consists of ppl who filled out wellbeing. 1 if friend, 0 if not. NA if no info.
  for (ii in 1:Len)
  {
    for (jj in 1:Len)
    {
        if (ii==jj)
        {
          w[ii,jj]=0
        }
      else
      {
        i=indexes[ii]
        j=indexes[jj]
        #print("i is ")
        #print(i)
        #print("j is ")
        #print(j)
        #print(df[i,j])
        #browser()
        if (df[i,j]==1)
        {
          w[ii,jj]=1
        }
        else
        {
          w[ii,jj]=0
        }
        
      }
      
    }
  }
  W<-sum(w)
  (x<-as.data.frame(res))
  numerator_s=0
  denominator_s=0
  x_mean=sum(x,na.rm=TRUE)/(dim(x)[1]-sum(rowSums(is.na(x)))) #correct. Checked manually 
  for (ii in 1:Len){
    x_i<-x[indexes[ii],1]
    for (jj in 1:Len){
      x_j<-x[indexes[jj],1]
      value=w[ii,jj]*(x_i-x_j)**2
      #print(value)
      numerator_s=numerator_s+value
    }
    denominator_s=denominator_s+(x_i-x_mean)**2
  }
  #print("numerator_s")
  #print(numerator_s)
  #print("denominator_s")
  #print(denominator_s)
  answer<-((N-1)*numerator_s)/ (2*W*denominator_s)
  answer
}

```

```{r}
k78<-wellbeing[,2]
k78<-k78[-m2]
length(k78)
dim(f1_1)
class(a)
class(b)
c<-as.matrix(k78)
class(c)
class(f1_1)
gearyC2(res=as.matrix(k78),meow=f1_1)
```