---
title: "RSiena goodness of fit"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r cars}
getwd()
```
```{r}
download.file("https://www.stats.ox.ac.uk/~snijders/siena/vdBunt_data.zip",
                                                        destfile='vdb.zip')
unzip('vdb.zip')
```

```{r}
vdb.w0 <- as.matrix(read.table("VRND32T0.DAT"))
vdb.w1 <- as.matrix(read.table("VRND32T1.DAT"))
vdb.w2 <- as.matrix(read.table("VRND32T2.DAT"))
vdb.w3 <- as.matrix(read.table("VRND32T3.DAT"))
vdb.w4 <- as.matrix(read.table("VRND32T4.DAT"))
vdb.w5 <- as.matrix(read.table("VRND32T5.DAT"))
vdb.w6 <- as.matrix(read.table("VRND32T6.DAT"))
vdb.attr <- as.matrix(read.table("VARS.DAT"))
```

```{r}
# Take account of missing data codes 6 and 9:
vdb.w0[vdb.w0 %in% c(6,9)] <- NA
vdb.w1[vdb.w1 %in% c(6,9)] <- NA
vdb.w2[vdb.w2 %in% c(6,9)] <- NA
vdb.w3[vdb.w3 %in% c(6,9)] <- NA
vdb.w4[vdb.w4 %in% c(6,9)] <- NA
vdb.w5[vdb.w5 %in% c(6,9)] <- NA
vdb.w6[vdb.w6 %in% c(6,9)] <- NA
```

```{r}
# Recode 4 (acquaintance) and 5 (difficult) to no tie
vdb.w0[vdb.w0 %in% c(4,5)] <- 0
vdb.w1[vdb.w1 %in% c(4,5)] <- 0
vdb.w2[vdb.w2 %in% c(4,5)] <- 0
vdb.w3[vdb.w3 %in% c(4,5)] <- 0
vdb.w4[vdb.w4 %in% c(4,5)] <- 0
vdb.w5[vdb.w5 %in% c(4,5)] <- 0
vdb.w6[vdb.w6 %in% c(4,5)] <- 0
```

```{r}
# Use the "friendly relation" relation by recoding:
vdb.w0[vdb.w0 %in% c(1,2,3)] <- 1
vdb.w1[vdb.w1 %in% c(1,2,3)] <- 1
vdb.w2[vdb.w2 %in% c(1,2,3)] <- 1
vdb.w3[vdb.w3 %in% c(1,2,3)] <- 1
vdb.w4[vdb.w4 %in% c(1,2,3)] <- 1
vdb.w5[vdb.w5 %in% c(1,2,3)] <- 1
vdb.w6[vdb.w6 %in% c(1,2,3)] <- 1
```

```{r}
library(RSiena)
# Attributes:
sex     <- coCovar(vdb.attr[,1])
program <- coCovar(vdb.attr[,2])
smoke   <- coCovar(vdb.attr[,3])
```

```{r}
# waves 2-3-4
friends234 <- sienaDependent(array(c(vdb.w2,vdb.w3,vdb.w4),
                            dim=c(32, 32, 3)), allowOnly=FALSE)
vdb.data234 <- sienaDataCreate(friends234,sex,program,smoke)
```

```{r}
# Algorithm creation:
vdb.algo <- sienaAlgorithmCreate(projname = 'vdb_gof234', seed=123456)
```

```{r}
?'sienaGOF-auxiliary'
#Auxiliary functions for goodness of fit assessment by sienaGOF
#The auxiliary functions are, first, some functions of networks or behavior (i.e., statistics) for which the simulated values for the fitted model are compared to the observed value
```

```{r}
#geodesic distance between two vertices is the length in terms of the number of edges of the shortest path between the vertices.

# GeodesicDistribution calculates the distribution of non-directed
   # geodesic distances
   # The default for \code{levls} reflects the usual phenomenon
   # that geodesic distances larger than 5
   # do not differ appreciably with respect to interpretation.
   # Note that the levels of the result are named;
   # these names are used in the \code{plot} method.

GeodesicDistribution <- function (i, data, sims, period, groupName,
   varName, levls=c(1:5,Inf), cumulative=TRUE, ...) {
     x <- networkExtraction(i, data, sims, period, groupName, varName)
     require(sna)
     a <- sna::geodist(symmetrize(x))$gdist
     if (cumulative)
     {
       gdi <- sapply(levls, function(i){ sum(a<=i) })
     }
     else
     {
       gdi <- sapply(levls, function(i){ sum(a==i) })
     }
     names(gdi) <- as.character(levls)
     gdi
}
```

```{r}
# The following function is taken from the help page for sienaTest

testall <- function(ans){
    for (i in which(ans$test))
      {
    sct <- score.Test(ans,i)
    cat(ans$requestedEffects$effectName[i], '\n')
    print(sct)
    }
    invisible(score.Test(ans))
}
```

Now,let's applying sienaGOF to the van de Bunt data for waves 2, 3, 4.
```{r}
# Define a model to decide between three specifications of transitivity.
# This forward model selection is not recommended in general;
# it is used here only to demonstrate the use of the sienaGOF function.

vdb.eff234 <- getEffects(vdb.data234)
vdb.eff234 <- includeEffects(vdb.eff234, transTrip, transTies, gwespFF,
                    test=TRUE, fix=TRUE)
vdb.eff234
```

```{r}
(ans0 <- siena07(vdb.algo, data=vdb.data234, effects=vdb.eff234,
                                                    returnDeps=TRUE))
```

The overall maximum convergence ratio is less than 0.25, so we don't repeat the algorithm. 

```{r}
testall(ans0)
```

```{r}
(gofi0 <- sienaGOF(ans0, IndegreeDistribution, verbose=TRUE, join=TRUE,
     varName="friends234"))
```

Q: Not sure how to interpret Monte Carlo p-value and calculated joint MHD
Q: Does verbose=True print all the messages

Q: what does join=True do? 
Ans: Note that join=FALSE is used to get separate calculations for the two periods; join=TRUE will calculate the sum over the periods.

```{r}
(gofo0 <- sienaGOF(ans0, OutdegreeDistribution, verbose=TRUE, join=TRUE,
     levls=c(0:10,15,20),varName="friends234"))
```

```{r}
(gof0.tc <- sienaGOF(ans0, TriadCensus, verbose=TRUE, join=TRUE,
     varName="friends234"))
(gof0.gd <- sienaGOF(ans0, GeodesicDistribution, cumulative=FALSE,
     verbose=TRUE, join=TRUE, varName="friends234"))
```

 Fit for indegree distribution is OK as p-value is greater than 0.05;for the other three auxiliary statistics, it is not as the p-value <0.05.
 
```{r}
plot(gof0.gd)
```
 #How to interpret this diagram? Similar question for the paper I read.
 
```{r}
# The model simulates too many geodesic distances 2, and too few distances 5
descriptives.sienaGOF(gof0.gd)
```
 
 Q: Do the column numbers represent geodesic distances? If so, how can Inf be the case?
 
```{r}
# What about distances 6 or 7?
(gof0.gd7 <- sienaGOF(ans0, GeodesicDistribution, cumulative=FALSE,
            levls=c(1:7,Inf), verbose=TRUE, join=TRUE, varName="friends234"))
descriptives.sienaGOF(gof0.gd7)
```
Q: What does cumulative=False do?
Q: From the website, these comments below are provided
# No distances 6 are observed, and some are simulated (depending on the random simulations)
# No distances 7 are observed, and none are simulated:
 In this case, I feel no distances 7 are observed, whereas distances 6 are observed. Check if this right? Is this due to the matter of seed and R version? I feel my belief is evident in code output below but still double check.
 
```{r}
descriptives.sienaGOF(gof0.gd7, showAll=TRUE)
```
 
```{r}
# The summaries give the approximate value of the Mahalanobis distance,
# if the tested & fixed effects would be added:
summary(gof0.tc)
summary(gof0.gd)
```
 Q: Interpreting one-step values
 Q: what do tc in gof0.tc and gd in gof0.gd stand for? 
 A:Triad census and geodesicDistribution
 
The summaries for gof0.tc and gof0.gd suggest that the model should be extended by gwespFF or transTrip. We go for gwespFF.
Q:How did we decide it should be extended by either and why did we go with gwespFF?

```{r}
vdb.eff234.a <- includeEffects(vdb.eff234, gwespFF)
(ansa <- siena07(vdb.algo, data=vdb.data234, effects=vdb.eff234.a,
                prevAns=ans0, returnDeps=TRUE))
testall(ansa)
# The two other transitivity effects now have become non-significant.
```
Q: How to use the result of testall? Since the p-values are all greater than 0.05, do we assume transitive triplets and ties are not statistically significant in the evolution of this network?

```{r}
(gofia <- sienaGOF(ansa, IndegreeDistribution, verbose=TRUE, join=TRUE,
     varName="friends234"))
(gofoa <- sienaGOF(ansa, OutdegreeDistribution, verbose=TRUE, join=TRUE,
     levls=c(0:10,15,20),varName="friends234"))
(gofa.tc <- sienaGOF(ansa, TriadCensus, verbose=TRUE, join=TRUE,
     varName="friends234"))
(gofa.gd <- sienaGOF(ansa, GeodesicDistribution, cumulative=FALSE,
     verbose=TRUE, join=TRUE, varName="friends234"))
```

Now the fit is OK for all four auxiliary statistics.
```{r}
plot(gofoa)
```
#Q: Same interpretation issue as before

To get a plot for the triad census, the values should be centered and scaled,because they differ wildly

```{r}
plot(gofa.tc)
plot(gofa.tc, center=TRUE, scale=TRUE)
```
Q: Are there other effects that need to be centered and scaled? Why does tc need this preprocessing?

```{r}
# Is the model homogeneous for the two periods?
sienaTimeTest(ansa)

```
Q: Check if my interpretation is correct
Answer is yes because the p-value for joint significance test of time heterogeneity was greater than 0.05. This means the model didn't change too much between the two periods

-----------

We want to get an impression of the distribution of the number of components. The components are computed by the function geo.desc through the package igraph.

igraphNetworkExtraction extracts simulated and observed networks from the results of a siena07 run.
It returns the network as an edge list of class "graph" according to the igraph package.

Q: I am not sure about the interpretation of explannation provided
"Ties for ordered pairs with a missing value for wave=period or period+1 are zeroed;
note that this also is done in RSiena for calculation of target statistics.
However, changing structurally fixed values are not taken into account."

```{r}
igraphNetworkExtraction <- function(i, data, sims, period, groupName, varName) {
  require(igraph)
  dimsOfDepVar<- attr(data[[groupName]]$depvars[[varName]], "netdims")
  missings <- is.na(data[[groupName]]$depvars[[varName]][,,period]) |
    is.na(data[[groupName]]$depvars[[varName]][,,period+1])
  if (is.null(i)) {
    # sienaGOF wants the observation:
    original <- data[[groupName]]$depvars[[varName]][,,period+1]
    original[missings] <- 0
    returnValue <- graph.adjacency(original)
  }
  else
  {
    missings <- graph.adjacency(missings)
    #sienaGOF wants the i-th simulation:
    returnValue <- graph.difference(
      graph.empty(dimsOfDepVar) +
            edges(t(sims[[i]][[groupName]][[varName]][[period]][,1:2])),
      missings)
  }
  returnValue
}
```

Now, let's calculate the size of the largest component,the number of components, and the diameter.
It uses igraph to calculate the components.

```{r}
geo.desc <- function(i, data, sims, period, groupName, varName){
    require(igraph)
    ai <- igraphNetworkExtraction(i, data, sims, period, groupName, varName)
    n  <- vcount(ai)  # number of nodes
    # size largest weak component
    componentList <- clusters(ai,mode="weak")
    comp1 <- max(componentList$csize)
    # number of components
    ncomp <- componentList$no
    # diameter
    diam <- diameter(as.undirected(ai), unconnected=TRUE)
    descr <- c(comp1, ncomp, diam)
    names(descr) <- c("comp1", "ncomp", "diam")
    descr
}
```

```{r}
# Note that join=FALSE is used here, to get separate calculations
# for the two periods; join=TRUE will calculate the sum over the periods.
(gofa.geo <- sienaGOF(ansa, geo.desc, verbose=TRUE, join=FALSE,
     varName="friends234"))
descriptives.sienaGOF(gofa.geo, period=1)
descriptives.sienaGOF(gofa.geo, period=2)
```

