---
title: "RSiena experiment 1"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r,message=FALSE}
tinytex::reinstall_tinytex()
library(RSiena)
library(igraph)
```

```{r}
friend.data.w1<-s501
friend.data.w2<-s502
friend.data.w3<-s503
drink<-s50a
smoke<-s50s
friendshipData <- array( c( friend.data.w1, friend.data.w2, friend.data.w3 ),
           dim = c( 50, 50, 3 ) )
friendship <- sienaDependent(friendshipData)
friendship
```

I just created a 50 * 50 * 3 array composed of the 3 adjacency matrices(friend datas).

Now, we use 1st wave of smoking as a constant actor covariate. Then, we define a variable actor covariate(alcohol) for drinking.Then,we put all the variables we've defined so far to create a dataset for analysis. 
```{r}
smoke1 <- coCovar( smoke[ , 1 ] )
alcohol <- varCovar( drink )
mydata <- sienaDataCreate( friendship, smoke1, alcohol )
mydata
#print01Report( mydata, modelname="s50")
```
Above contains some basic descriptives of the dataset. Also, I just created an output file s50.txt in your current working directory. Note that if an existing file has this name, it will be overwritten!

---

For the model specification, an effects object is needed, so I created one named myeff.
```{r}
myeff <- getEffects( mydata )
#effectsDocumentation(myeff)
```

```{r}
?includeEffects
myeff <- includeEffects( myeff, transTrip, cycle3)
myeff <- includeEffects( myeff, egoX, altX, simX, interaction1 = "alcohol" )
#cuz we used all of alcohol when creating mydata

#>   effectName         include fix   test  initialValue parm
#> 1 alcohol alter      TRUE    FALSE FALSE          0   0   
#> 2 alcohol ego        TRUE    FALSE FALSE          0   0   
#> 3 alcohol similarity TRUE    FALSE FALSE          0   0
myeff <- includeEffects( myeff, simX, interaction1 = "smoke1" ) 
#cuz we used one constant covariate for smoke previously when making mydata

#>   effectName        include fix   test  initialValue parm
#> 1 smoke1 similarity TRUE    FALSE FALSE          0   0
#myeff
```

transTrip
- exploring the effect of the existence of ties i --> h and h --> j on the formation of tie i --> j.

cycle3
-number of three-cycles: If there is a connection between i,j,h such that there is a tie between  x_ij,x_jh,x_hi, then there is one three-cycle.

(egoX, altX, simX, interaction1 = "alcohol")
(simX, interaction1 = "smoke1")

*egoX,altX: For centered actor covariates, function sienaGroupCreate centers the original values
again, now around the overall mean. This allows the use of centered actor covariates that
are constant within groups. The egoX effect of a centered covariate will be the sender
effect of the original variable with grand-mean centering. For the altX effect the centering
makes no difference
*simX: For homophily parameters on important attributes expressed by the simX effect (which is
standardized).
#Homophily: the tendency for people to seek out or be attracted to those who are similar to themselves.

#Q: I am still not sure about importance of altX. Also, what would be centered covariate in this case? 

I now will create an object with algorithm settings. To ensure it produces same estimation result consistently, I used seed parameter. 
```{r}
?sienaAlgorithmCreate
myalgorithm <- sienaAlgorithmCreate( projname = 's50',seed=435123  )
```

```{r}
ans <- siena07( myalgorithm, data = mydata, effects = myeff)
ans
```
For checking convergence, I looked at the overall maximum convergence ratio. As it is less than 0.25, convergence is good. Otherwise, I would've estiamted once more, using the result obtained as the previous ans from which estimation continues.
ex. siena07( myalgorithm, data = mydata, effects = myeff, prevAns=ans)   

---

Networks and Behavior study

For this analysis, we redefine the role of alcohol drinking as a dependent behaviour variable via 'type' parameter.
```{r}
drinking <- sienaDependent( drink, type = "behavior" )
NBdata <- sienaDataCreate( friendship, smoke1, drinking )
#NBdata
```

Now, we create a variable named NBeff, which represents bundle of effects.

transTrip (previously explained)

transRecTrip :transitive reciprocated triplets effect (transRecTrip), which can be regarded as an interaction between the transitive triplets effect and reciprocity, where the reciprocated tie is the tie i <--> j that closes
the two-path i->h->j.

egoX (previously explained)

egoSqX: covariate squared - ego or squared covariate-related activity (egoSqX), defined by i out-degree weighted by his covariate value.

altX (previously explained)
altSqX:  covariate squared - alter or squared covariate-related popularity (altSqX), defined by the sum of the squared centered covariate over all actors to whom i has a tie, (not included if the variable has range less than 2)

diffSqX: covariate-squared-difference or covariate-related squared difference (diffSqX), defined by the squared alter-minus-ego difference of the covariate over all actors to whom i has a tie.

simX (previously explained)

avAlt: average alter effect (avAlt), defined by i’s behavior multiplied by the average behavior of his alters (a kind of ego-alter behavior covariance),

#Q or comment: Personally, I am just testing out different effects; I am not sure whether all of them are relevant or helpful.

```{r}
NBeff <- getEffects( NBdata )
effectsDocumentation(NBeff)
NBeff <- includeEffects( NBeff, transTrip, transRecTrip )
#>   effectName                  include fix   test  initialValue parm
#> 1 transitive triplets         TRUE    FALSE FALSE          0   0   
#> 2 transitive recipr. triplets TRUE    FALSE FALSE          0   0
NBeff <- includeEffects( NBeff, egoX, egoSqX, altX, altSqX, diffSqX,
                         interaction1 = "drinking" )
#>   effectName             include fix   test  initialValue parm
#> 1 drinking alter         TRUE    FALSE FALSE          0   0   
#> 2 drinking squared alter TRUE    FALSE FALSE          0   0   
#> 3 drinking ego           TRUE    FALSE FALSE          0   0   
#> 4 drinking squared ego   TRUE    FALSE FALSE          0   0   
#> 5 drinking diff. squared TRUE    FALSE FALSE          0   0
NBeff <- includeEffects( NBeff, egoX, altX, simX, interaction1 = "smoke1" )
#>   effectName        include fix   test  initialValue parm
#> 1 smoke1 alter      TRUE    FALSE FALSE          0   0   
#> 2 smoke1 ego        TRUE    FALSE FALSE          0   0   
#> 3 smoke1 similarity TRUE    FALSE FALSE          0   0
NBeff <- includeEffects( NBeff, avAlt, name="drinking",
                         interaction1 = "friendship" )
#We now have to mention the name of the dependent behavior variable; it did not need to be mentioned before, because the default name is the first in the data set, which is friendship.
#NBeff
```

```{r}
myalgorithm1 <- sienaAlgorithmCreate( projname = 's50_NB',seed=42)
NBans <- siena07( myalgorithm1, data = NBdata, effects = NBeff)
NBans
```

There is no need to do this process again since overall maximum convergence ratio is less than 0.25, but I am repeating it to see the effect. As one would expect, overall maximum convergence ratio decreased. 

#Q: What is the consequence/risk of repeating this process when the ratio is already less than 0.25?

```{r}
NBans1 <- siena07( myalgorithm1, data = NBdata, effects = NBeff,prevAns=NBans)
NBans1
```

Code chunk below produces a nicer listing of the results. 

```{r}
#siena.table(NBans, type="html", sig=TRUE)
```

---Feedback/learning:
egoX= If I drink more alcohol, am I more likely to send out more friendship ties
Alt:alter . Ppl drinking, are they more likely to receive ties? 
Smok1 and alcohol would be centred covariates 

Method of moment to estimate ratio. Finding parameters to where expected value=observed value. 
Cannot find closed form. 
Expression of difference between expected value and observed value 
Parameter estimates don’t change much even if you repeat after getting less than 0.25 so not worth. 


#Q or comment: Personally, I am just testing out different effects; I am not sure whether all of them are relevant or helpful.
Ans: look at data +RSiena manual for essential ones

Convergence ratio Smaller than 0.1, repeat algorithm 

rate: on average, actors in network get 6.31 opportunities to make changes 
eval: if smoke, more likely to be selected as a friend if convergence t-ratio<0.05

eval drinking ego: if drink, more likely to send friends. 
#send>make as it clarifies the direction. 

—
Co-evolution of network by prof. 

Look at method papers by ppl who developed RSiena

—
Away from pgh a lot this sem. Working remote. 
Next sem meet in person 

-read paper
-refer to goodness of fit
-refer to her code 
